Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-01-31 overview.tex
Switched prime and non-prime variables in overview.

\begin{eqnarray} 
\left( \frac{\partial \rhop}{\partial t} \right)_r 
          + \vecvp \cdot \nabla_r \rhop
    & = & - \rhop \nabla_r \cdot \vecvp,
        \label{eq:mass_prime} \\
%
\left( \frac{\partial \vecvp }{ \partial t} \right)_r 
          + ( \vecvp \cdot \nabla_r ) \vecvp 
    & = & - \frac{1}{\rhop} \nabla_r \pp 
          - \nabla_r \phip,
        \label{eq:momentum_prime} \\
%
\left( \frac{\partial \Ep}{\partial t} \right)_r + \vecvp \cdot \nabla_r \Ep
    & = & - \frac{1}{\rhop} \nabla_r \cdot (\pp \vecvp) 
          - \vecvp \cdot \nabla_r \phip
%\nonumber          \\  & &     % uncomment for 2-column
     + \Gamma - \Lambda
        \label{eq:total_energy_prime}
\end{eqnarray}
%
Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the baryonic total specific fluid energy, the baryonic specific thermal energy, the baryonic density, velocity and pressure, respectively.  Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.  They are respectively, the first, second and third moments of the Boltzman equation. 
Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy_prime}).

The total fluid specific energy ($\Ep$) is given by:
\begin{equation}
\Ep =  e^\prime + \frac{1}{2} \vecvp \cdot \vecvp.
        \label{eq:total_energy_def_prime}
\end{equation}
It is useful to solve the equation of total energy because doing so enforces local energy conservation.

The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
%
\begin{eqnarray}
%
e^\prime    & = & \pp / [(\gamma - 1) \rhop ],
        \label{eq:eq_of_state_prime} \\
%
\nabla_r^2 \phip & = & 4 \pi G \rho_{total}^\prime 
%                   + 3 p_{total}^\prime/c^2) - \Lambda. 
        \label{eq:potential_prime}
\end{eqnarray}
%
For simplicity, the equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
Equation (\ref{eq:potential_prime}) includes contributions to the gravitational potential $\phi^\prime$ from all  populations ($\rho_{total}^\prime$).
% as well as any radiation or massless neutrino background ($p_{total}^\prime$) or cosmological constant ($\Lambda_c$), if we are interested in the cosmological case.

@@

This set of coupled hydrodynamic plus gravity equations is the standard one that \enzo solves.  However, in the cosmological case it is useful to transform these to a coordinate system that is comoving with the expanding universe \citep{Peebles93} (but note that \enzo\ can also operate in a fixed coordinate system). Specifically, we define:
\begin{eqnarray}
\vecx & = & \vecr/a, \\
\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
              \vecvp - \dot{a}\vecx, \\
\rho    & = & a^3 \rhop,   \\
p       & = & a^3 \pp, \\
E       & = & \Ep - 
              \dot{a} \vecx \cdot \vecv - 
              \frac{1}{2} \dot{a}^2 \vecx^2, \\
\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2.
\end{eqnarray}

These definitions result in the following comoving equivalents to eqs. 
(\ref{eq:mass_prime})-(\ref{eq:potential_prime}).
%

\begin{eqnarray} 
\frac{\partial \rhob}{\partial t} 
          + \frac{1}{a} \vecvb \cdot \nabla \rhob 
    & = & - \frac{1}{a} \rhob \nabla \cdot \vecvb, 
        \label{eq:mass} \\
%
\frac{\partial \vecvb}{\partial t}  
          + \frac{1}{a} ( \vecvb \cdot \nabla ) \vecvb 
    & = & - \frac{\dot{a}}{a} \vecvb 
          - \frac{1}{a \rhob} \nabla p 
          - \frac{1}{a} \nabla \phi, 
        \label{eq:momentum} \\
%
\frac{\partial E}{\partial t} 
         + \frac{1}{a} \vecvb \cdot \nabla E 
   & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rhob} + \vecvb^2 )
         - \frac{1}{a \rhob } \nabla \cdot (p \vecvb) 
%\nonumber   & &   %uncomment for 2-column
         - \frac{1}{a} \vecvb \cdot \nabla \phi 
                + \Gamma - \Lambda.
       \label{eq:total_energy} \\
%
E  & = & e + \frac{1}{2} {\vecvb}^2,
        \label{eq:total_energy_def} \\
%
e                  & = & 
        p/\left[ \left( \gamma - 1 \right) \rhob \right],
      \label{eq:state} \\ 
{\nabla}^2 \phi    & = & 
        \frac{4\pi G}{a} ( \rho_{total} - \rho_0 ).
      \label{eq:poisson}
\end{eqnarray}
%
%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by 
%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
%The density, however, is comoving:
%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of a smooth, homogeneous background, where $z$, the redshift, is a function only of $t$.  All derivatives are determined with respect to the comoving position $\vecx$, which is defined simply to remove the universal expansion from the coordinate system.

The evolution of $a(t)$ is governed by the formula for the expansion of an isotropic, homogeneous universe:
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rho_0 
      + 3p_0/c^2) 
      + \Lambda_c /3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here, $\rho_0$ is the mean (baryonic and dark) comoving mass density and $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and assumes that curvature effects are not important --- both assumptions are reasonable as long as the  size of the simulated region is small compared to the radius of curvature and the Hubble length $c/H$ ($c$ is the speed of light and $H$ is the Hubble constant).

@@

\begin{equation}
\mathbb{F} = -\kappa_{sp} \^{b} (\^{b} \cdot \nabla T)
\end{equation}

Where $\^{b}$ is the unit vector denoting the direction of the
magnetic field.  In both the isotropic and anisotropic limits, we
assume that conductivity can be suppressed relative to the ideal
Spitzer conductivity, and express this as a constant multiplier.

>>>>

\begin{eqnarray} 
\frac{\partial \rho}{\partial t} 
          + \vecv \cdot \nabla \rho
    & = & - \rho \nabla \cdot \vecv,
        \label{eq:mass} \\
%
\frac{\partial \vecv }{ \partial t} 
          + ( \vecv \cdot \nabla ) \vecv 
    & = & - \frac{1}{\rho} \nabla p
          - \nabla \phi,
        \label{eq:momentum} \\
%
 \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
          - \vecv \cdot \nabla \phi
%\nonumber          \\  & &     % uncomment for 2-column
     + \Gamma - \Lambda
        \label{eq:total_energy}
\end{eqnarray}
%
Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the baryonic total specific fluid energy, the baryonic specific thermal energy, the baryonic density, velocity and pressure, respectively.  
%Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.  They are respectively, the first, second and third moments of the Boltzmann equation. 
Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy_prime}).

The total fluid specific energy ($E$) is given by:
\begin{equation}
E =  e + \frac{1}{2} \vecv \cdot \vecv.
        \label{eq:total_energy_def}
\end{equation}
It is useful to solve the equation of total energy because doing so enforces local energy conservation.

The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
%
\begin{eqnarray}
%
e   & = & p / [(\gamma - 1) \rho ],
        \label{eq:eq_of_state} \\
%
\nabla^2 \phi & = & 4 \pi G \rho_{total}
%                   + 3 p_{total}/c^2) - \Lambda. 
        \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
Equation (\ref{eq:potential}) includes contributions to the gravitational potential $\phi$ from the total mass density $\rho_{\rm total}$).
% as well as any radiation or massless neutrino background ($p_{total}^\prime$) or cosmological constant ($\Lambda_c$), if we are interested in the cosmological case.

@@

This set of coupled hydrodynamic plus gravity equations is the standard one that \enzo solves.  However, in the cosmological case it is useful to transform these to a coordinate system that is comoving with the expanding universe \citep{Peebles93} (but note that \enzo\ can also operate in a fixed coordinate system). Specifically, we define:
\begin{eqnarray}
\vecx & = & \vecr/a, \\
\vecvp & = & a \hspace{0.5mm} d\vecx/dt = 
              \vecv - \dot{a}\vecx, \\
\rhop    & = & a^3 \rho,   \\
\pp       & = & a^3 p, \\
\Ep       & = & E - 
              \dot{a} \vecx \cdot \vecvp - 
              \frac{1}{2} \dot{a}^2 \vecx^2, \\
\phip    & = & \phi + \frac{1}{2} a \ddot{a} \vecx^2.
\end{eqnarray}

These definitions result in the following comoving equivalents to eqs. 
(\ref{eq:mass})-(\ref{eq:potential}).
%

\begin{eqnarray} 
\frac{\partial \rhop}{\partial t} 
          + \frac{1}{a} \vecvp \cdot \nabla \rhop 
    & = & - \frac{1}{a} \rhop \nabla \cdot \vecvp, 
        \label{eq:mass_prime} \\
%
\frac{\partial \vecvp}{\partial t}  
          + \frac{1}{a} ( \vecvp \cdot \nabla ) \vecvp 
    & = & - \frac{\dot{a}}{a} \vecvp 
          - \frac{1}{a \rhop} \nabla \pp 
          - \frac{1}{a} \nabla \phip, 
        \label{eq:momentum_prime} \\
%
\frac{\partial \Ep}{\partial t} 
         + \frac{1}{a} \vecvp \cdot \nabla \Ep 
   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
         - \frac{1}{a \rhop } \nabla \cdot (p \vecvp) 
%\nonumber   & &   %uncomment for 2-column
         - \frac{1}{a} \vecvp \cdot \nabla \phip 
                + \Gamma - \Lambda.
       \label{eq:total_energy_prime} \\
%
\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
        \label{eq:total_energy_def_prime} \\
%
\ep                  & = & 
        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
      \label{eq:state_prime} \\ 
{\nabla}^2 \phip    & = & 
        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
      \label{eq:poisson}
\end{eqnarray}
%
%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by 
%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
%The density, however, is comoving:
%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of a smooth, homogeneous background, where $z$, the redshift, is a function only of $t$.  All derivatives are determined with respect to the comoving position $\vecx$, which is defined simply to remove the universal expansion from the coordinate system.

The evolution of $a(t)$ is governed by the formula for the expansion of an isotropic, homogeneous universe:
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rhop_0 
      + 3\pp_0/c^2) 
      + \Lambda_c /3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here, $\rhop_0$ is the mean (baryonic and dark) comoving mass density and $\pp_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and assumes that curvature effects are not important --- both assumptions are reasonable as long as the  size of the simulated region is small compared to the radius of curvature and the Hubble length $c/H$ ($c$ is the speed of light and $H$ is the Hubble constant).

@@

\begin{equation}
\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
\end{equation}

Where $\hat{b}$ is the unit vector denoting the direction of the
magnetic field.  In both the isotropic and anisotropic limits, we
assume that conductivity can be suppressed relative to the ideal
Spitzer conductivity, and express this as a constant multiplier.
